\vspace{-0.25em}\thetitle

### 2009-07-29 paper.tex
...

\begin{abstract}
We have observed that when network traffic behaviors are represented in vector spaces as relative frequency histograms of behavioral features, they tend to exhibit a great deal of low-rank linear structure.
We hypothesize that this structure is due the distribution of flow behaviors following a  finite mixture model.
Aside from being of theoretical interest, this hypothesis has practical consequences:
it allows us, among other things, to make detailed predictions about the probabilities of various future flow behaviors after observing only a handful of a flow's initial packets.
This practical application serves a dual function.
It provides a highly useful tool of network management, especially in wireless networks where such predictions can be leveraged to better allocate scarce bandwidth resources.
However, it also provides evidence that the hypothesized model provides a correct explanation for the observed linear structure in real network traffic.
\end{abstract}

@@

% It seems wise at this point to address the nature of prediction for highly nondeterministic phenomena like network traffic.
Brief reflection shows that one cannot hope to accurately predict the exact behavior of flows:
flows may exhibit the same initial behavior but subsequently behave quite differently.
The best we can hope to do is to predict a \emph{distribution} of possible outcomes given the observed initial behavior.
When compared with a specific outcome, such a prediction cannot be said to be unequivocally right or wrong.
How then can we measure the quality of a predicted distribution of outcomes as compared to a single outcome?
This subtle and difficult issue is central to our evaluation methodology.

@@

>>>>

\begin{abstract}
We have observed that when network traffic behaviors are represented in vector spaces as relative frequency histograms of behavioral features, they tend to exhibit a great deal of low-rank linear structure.
We hypothesize that this structure is due the distribution of flow behaviors following a  finite mixture model.
Aside from being of theoretical interest, this hypothesis has practical consequences:
it allows us, among other things, to make detailed, accurate predictions about the probabilities of various future flow behaviors after observing only a handful of a flow's initial packets.
% We are able to predict the distribution of future packet sizes and inter-packet intervals from only five initial packets with between 70\% and 90\% accuracy across a variety of network traces.
% For half of flows, we can reduce the uncertainty of our packet count predictions by 80\% or more from the same observations.
This practical application serves a dual function.
It provides a highly useful tool of network management, especially in wireless networks where such predictions can be leveraged to better allocate scarce bandwidth resources.
However, it also provides evidence that the hypothesized model provides a correct explanation for the observed linear structure in real network traffic.
\end{abstract}

@@

% It seems wise at this point to address the nature of prediction for highly nondeterministic phenomena like network traffic.
Brief reflection shows that one cannot hope to accurately predict the exact behavior of flows:
flows may exhibit the same initial behavior but subsequently behave quite differently.
The best we can hope to do is to predict a \emph{distribution} of possible outcomes given the observed initial behavior.
When compared with a specific outcome, such a prediction cannot be said to be unequivocally right or wrong.
How then can we measure the quality of a predicted distribution of outcomes as compared to a single outcome?
This subtle and difficult issue is central to our evaluation methodology.
% In what follows, we have performed prediction from observing only five initial packets from each flow.

@@

Our results show that we can predict the packet size distribution and inter-packet interval distribution of flows from only observing a handful of initial packets of each flow.
Specifically, we have used five initial packets to predict these distributions with between 70\% and 90\% accuracy across a wide variety of network traces.
Moreover, we can reduce our uncertainty about the number of packets that flows will have by 80\% or more, about half of the time.
Immediate applications of this ability to accurately predict individual flow behaviors abound.
Quality of service (\caps{QoS}) schemes attempting to provide high-quality service guarantees for streaming voice and video wireless through wireless networks could use such predictions to know---almost presciently---how much data flows will transmit and how often.

The real importance of this work has little to do with these potential practical applications.
The ability to predict flow behavior is merely a pleasant side effect of the real breakthrough:
a detailed statistical model for network-wide flow behaviors.
The fundamental concept underlying this work is that there are a relatively small collection of ``basic behaviors'' that generate the vast majority of the flow behaviors that we actually see in any particular trace.
These behaviors are associated with distributions of various flow features:
packet sizes, inter-packet intervals, \caps{IP} protocol, port numbers and packet counts.
these features are drawn from these distributions to ``create'' actual flows.

% TODO: mention the two dimension structure assumption.
% TODO: mention the five packets in the results somewhere.
